---
title: ""
output: semantic.doc::semantic_document
---


```{r, warning=F, message=F, error=F, echo = F}
pacman::p_load(blogdown, shiny, shiny.semantic, dplyr, purrr)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval = T, message = NA, error = F, warning = F)
```


<style>
.myrow {
    margin-bottom: 0px !important;
    margin-left: 0px !important;
    margin-right: 0px !important;
    margin-top: 0px !important;
    margin: 0px;
    padding-bottom: 0px !important;
    padding-left: 0px !important;
    padding-right: 0px !important;
    padding-top: 0px !important;
    padding: 0px;
}

.ui.header {
  margin-left: 0px !important;
  margin-right: 0px !important;
  margin-bottom: 10px !important;
}
</style>

<br>

<div class = "ui container">

<!-- <a class="ui blue big right corner label" href = "http://www.systats.eu/cv_print/cv_simon_roth.pdf" target="_blank"> -->
<!--   <i class="file pdf icon"></i> -->
<!-- </a> -->

<a class="ui massive corner label" href="/cv_print/cv_simon_roth.pdf" target="_blank">
  <i class="file pdf icon"></i>
</a>


<h1 class = "ui massive centered header">
Simon Roth 
</h1>


<br>

```{r}
pacman::p_load(shiny, shiny.semantic, dplyr, purrr)

cv_entries <- function(x){
  x %>% 
    split(1:nrow(.)) %>% 
    purrr::map(~{
      div(class = "myrow row",
        div(class = "right aligned four wide column",
          div(class = "item", style ="margin-bottom: 5px",
            shiny::HTML(.x$short)    
          )
        ),
        div(class = "thirdten wide column",
          div(class = "item",  style ="margin-bottom: 5px",
            shiny::HTML(.x$long)
          )
        )
      )
    }) %>% 
    div(class = "ui divided grid", .) %>% 
    tagList(.)
}

cv_entries_desc <- function(x){
  x %>% 
    split(1:nrow(.)) %>% 
    purrr::map(~{
      div(class = "myrow row",
        div(class = "right aligned four wide column",
          div(class = "item", style ="margin-bottom: 5px",
            .x$short
          )
        ),
        div(class = "thirdten wide column",
          div(class = "item", style ="margin-bottom: 5px",
            shiny::HTML(.x$long), 
            if(.x$meta != "") tagList(br(), shiny::HTML(.x$meta))
          )
        )
      )
    }) %>% 
    div(class = "ui divided grid", .) %>% 
    tagList(.)
}

cv_icon_entries <- function(x){
  x %>% 
    split(1:nrow(.)) %>% 
    purrr::map(~{
      div(class = "myrow row",
        div(class = "right aligned four wide column",
          shiny::HTML(.x$short)
        ),
        div(class = "thirdten wide column",
          shiny::HTML(.x$long)
        )
      )
    }) %>% 
    div(class = "ui divided grid", .) %>% 
    tagList(.)
}

rbind(
  c("<i class='building icon'></i>", "University Konstanz, Germany"),
  c("<i class='map pin icon'></i>", "Room E 220"), 
  c("<i class='phone icon'></i>", "+49 7531-88-4127"),
  c("<i class='envelope outline icon'></i>", "<a href = 'mailto:s.roth@uni-konstanz.de'>s.roth@uni-konstanz.de</a>"),
  c("<i class='globe icon'></i>", 'Blog <i class ="right chevron icon"></i> <a href = "http://www.systats.eu/" target="_blank">systats.eu</a>'),
  c("<i class='github icon'></i>", 'Github <i class="right chevron icon"></i> <a href = "https://github.com/systats" target="_blank">systats</a>')
) %>%
  as_tibble %>% 
  set_names(c("short", "long")) %>% 
  cv_icon_entries
```



<br>
<br>


```{r}
rbind(
  c("", '<div class = "ui header"> Education </div>', ""),
  c("since 2018", "PhD Student at the Graduate School of Decision Sciences (<a href = 'https://www.gsds.uni-konstanz.de/' target='_blank'>GSDS</a>) and member of the Center for Data and Methods (<a href = 'https://www.polver.uni-konstanz.de/cdm/' target='_blank'>CDM</a>), University of Konstanz. Supervisor: <i>Susumu Shikano</i>", "<a class='ui gray horizontal label'>Major</a> Information Processing and Statistical Analysis <br> <a class='ui gray horizontal label'> Minor </a>
 Political Decisions and Institutions"), 
  c("2015 - 2018", "Master’s Degree in Empirical Research in Social Sciences at Stuttgart University. Supervisior: <i>André Bächtiger</i>", "<a class='ui gray horizontal label'> Thesis </a> The Good, the Bad, the Ugly: The strategic use of sentiment in parliamentary speeches."),
  c("2012 - 2015", "Bachelor’s Degree in Social Sciences at Stuttgart University", ""),
  c("1999 - 2012", "Abitur at Freie Waldorfschule Vaihingen/Enz.", "")
) %>%
  as_tibble %>% 
  set_names(c("short", "long", "meta")) %>% 
  cv_entries_desc
```


<br>
<br>


```{r}
rbind(
  c("", '<div class = "ui header"> Jobs & Academia </div>'),
  c("since 2017", "Data Scientist at <a href = 'https://www.paraboost.de/' target='_blank'>Paraboost </a>GmBH and at <a href = 'http://www.lov.ai/' target='_blank'>lov.ai <i class ='red heart outline icon'></i>  </a>"),
  c("2015 - 2017", "Research assistant within several NCCR projects on democratic preferences and political attitudes measured with experimental survey designs (Supervision André Bächtiger)."),
  c("2015 - 2016", "Teaching assistant for the lecture International Relations and European Integration"),
  c("Sep - Nov 2014", "Internship at Leibniz-Institut for Social Sciences (GESIS Mannheim): Department for Survey Design and Methodology.")
) %>%
  as_tibble %>% 
  set_names(c("short", "long")) %>% 
  cv_entries
```


<br>
<br>


```{r}
rbind(
  # c("2018", "Decoding the Alt-Right - Machine Learning and NLP project with a presnetation at VOX-Pol Amsterdam."),
  # c("2016", "ECPR General Conference, Prague with a paper presentation: A Deliberative Experiment on Democratic Preferences: Evaluating Preference Formation and Change.")
  c("", '<div class = "ui header"> Workshops & Conferences </div>
'),
  c("Aug 2018", "VOX-Pol Conference Amsterdam with a presentation: Decoding the Alt-Right"),
  c("May 2018", "Combating Online Extremism, VOX-Pol Network of Excellence, VOX-Pol Workshop at University of Oxford."),
  c("Jun 2018", "Challenges of Terrorism Informatics, VOX-Pol Network of Excellence,
VOX-Pol Workshop at Dublin City University."),
  c("Jun 2017", "Participation at UseR! 2017 in Brussels."), 
  c("Aug 2016", "ECPR Method Summer School, Budapest. Workshop on Bayesian Statistics.")
) %>%
  as_tibble %>% 
  set_names(c("short", "long")) %>% 
  cv_entries
```


<!-- <div class = "ui header"> -->
<!--   Working Papers -->
<!-- </div> -->

<!-- <div class = "ui header"> -->
<!--   Research Interest -->
<!-- </div> -->

```{r, eval = F}
rbind(
  c("substantive", "Collective Behaviour, Online Extremesim & Hate Speech"),
  c("methods", "NLP, machine learning, Bayesian statistics")
) %>%
  as_tibble %>% 
  set_names(c("short", "long")) %>% 
  cv_entries
```


<br>
<br>


<div class = "ui divided grid">
  <div class = "myrow row">
  <div class = "right aligned four wide column">
  </div>
  <div class = "thirdten wide column">
  <div class = "ui header"> Technical Skills </div>
  </div>
 </div>
  <div class = "myrow row">
  <div class = "right aligned four wide column">
  Back-End
  </div>
  <div class = "thirdten wide column">
  everyday R, sometimes Python & Unix
  </div>
 </div>
  <div class = "myrow row">
  <div class = "right aligned four wide column">
  ML
  </div>
  <div class = "thirdten wide column">
  Keras, H2O, SpaCy
  </div>
</div>
  <div class = "myrow row">
  <div class = "right aligned four wide column">
  Front-End
  </div>
  <div class = "thirdten wide column">
  Shiny Apps and Semantic UI
  </div>
</div>
  <div class = "myrow row">
  <div class = "right aligned four wide column">
  Reports
  </div>
  <div class = "thirdten wide column">
  
  
  Markdown, Rmarkdown, $\LaTeX$
  
  
 </div>
 </div>
</div>



<br>
<br>
<br>

<div align = "right">
  Last Update: `r Sys.Date()`
</div>

<br>
<br>
<br>

<!-- close container -->
</div>
